\modLP{Identification of} Biological Regulatory Networks from Process Hitting models

### 2014-09-15 parts-ph.tex
PH section: rel w/ Petri net + update cooperation paragraph

\todo{(LP) rework text, make presentation more CS oriented}

@@

\begin{example}
\todo{\pref{fig:runningPH-1} lacks readability (arrows hitting $a$ are too small).}
\pref{fig:runningPH-1} represents a PH $(\PHs,\PHl,\PHa)$ with
$\PHs = \{a,b,c\}$,
$\PHl_a = \{a_0,a_1,a_2\}$,
$\PHl_b = \{b_0, b_1\}$,
$\PHl_c = \{c_0, c_1\}$, and
\begin{align*}
\PHa & = \{
	\PHfrappe{a_2}{b_1}{b_0},
&&  \PHfrappe{b_0}{a_2}{a_1},
&&	\PHfrappe{c_0}{a_2}{a_1},\\
&&& \PHfrappe{b_0}{a_1}{a_0},
&&	\PHfrappe{c_0}{a_1}{a_0},\\
&&& \PHfrappe{b_1}{a_0}{a_1},
&&	\PHfrappe{c_1}{a_0}{a_1},\\
&&& \PHfrappe{b_1}{a_1}{a_2},
&&	\PHfrappe{c_1}{a_1}{a_2} \}\enspace.
\end{align*}
The action $h=\PHfrappe{b_1}{a_1}{a_2}$ is playable in the state
$s = \state{b_1,a_1,c_0}$; and $s\play h=\state{b_1,a_2,c_0}$.

@@

\paragraph{Modeling cooperation}
As described in \cite{PMR10-TCSB}, the cooperation between several processes to make another process bounce can be
expressed in PH by building a \emph{cooperative sort}.
\pref{fig:PH-cooperativity} shows an example of cooperation between processes $b_1$ and $c_1$ to
make $a_1$ bounce to $a_2$:
a cooperative sort $bc$ is defined with 4 processes (one for each sub-state of the presence of
processes $b_1$ and $c_1$).
For the sake of clarity, the processes of $bc$ are indexed using the sub-state they represent.
Hence, $bc_{01}$ represents the sub-state $\state{b_0,c_1}$, and so on.
Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the status of the sorts $b$
and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
Then, it is process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of
independent hits from $b_1$ and $c_1$.

@@

We note that cooperative sorts are standard PH sorts and do not involve any
special treatment regarding the semantics of related actions.

>>>>

@@

\modLP{PH can be seens as a particular sub-class of safe Petri nets \cite{BC92},
i.e. Petri nets with at most one token per place, having
groups of multually exclusive places (processes) acting as the sorts, and where each
transition has at most one read-arc (connected to the \emph{hitter}), and
exactly one incoming arc (from the \emph{target}) and one outgoing arc (to the
\emph{bounce}).
The relationship between PH and Petri nets is detailed in \cite{TR-PHvsPN-12}.}

\begin{example}
\pref{fig:runningPH-1} represents a PH $(\PHs,\PHl,\PHa)$ with
$\PHs = \{a,b,c\}$,
$\PHl_a = \{a_0,a_1,a_2\}$,
$\PHl_b = \{b_0, b_1\}$,
$\PHl_c = \{c_0, c_1\}$, and
\begin{align*}
\PHa & = \{
	\PHfrappe{a_2}{b_1}{b_0},
&&  \PHfrappe{b_0}{a_2}{a_1},
&&	\PHfrappe{c_0}{a_2}{a_1},\\
&&& \PHfrappe{b_0}{a_1}{a_0},
&&	\PHfrappe{c_0}{a_1}{a_0},\\
&&& \PHfrappe{b_1}{a_0}{a_1},
&&	\PHfrappe{c_1}{a_0}{a_1},\\
&&& \PHfrappe{b_1}{a_1}{a_2},
&&	\PHfrappe{c_1}{a_1}{a_2} \}\enspace.
\end{align*}
The action $h=\PHfrappe{b_1}{a_1}{a_2}$ is playable in the state
$s = \state{b_1,a_1,c_0}$; and $s\play h=\state{b_1,a_2,c_0}$.

@@

\paragraph{Modeling conjunctions}
\modLP{PH restricts the causality of transitions within sorts to the presence of
at most one process in another sort (the hitter).
One may want to model transitions that should be taken only when several
processes are present, i.e., modeling conjunctions -- or cooperations -- between the presence of
several processes.
As described in \cite{PMR10-TCSB}, such cooperations can be encoded in PH with
the use of some intermediate sorts, so-called \emph{cooperative sorts}, that
merge information on the state of cooperating sorts.}
\pref{fig:PH-cooperativity} shows an example of cooperation between processes $b_1$ and $c_1$ to
make $a_1$ bounce to $a_2$:
a cooperative sort $bc$ is defined with 4 processes (one for each sub-state of the presence of
processes $b_1$ and $c_1$).
For the sake of clarity, the processes of $bc$ are indexed using the sub-state they represent.
Hence, $bc_{01}$ represents the sub-state $\state{b_0,c_1}$, and so on.
Each process of sort $b$ and $c$ hit $bc$ to make it bounce to the process reflecting the status of the sorts $b$
and $c$ (e.g., $\PHfrappe{b_1}{bc_{00}}{bc_{10}}$ and $\PHfrappe{b_1}{bc_{01}}{bc_{11}}$).
Then, it is process $bc_{11}$ which hits $a_1$ to make it bounce to $a_2$ instead of
independent hits from $b_1$ and $c_1$.

@@

We note that cooperative sorts are standard PH sorts and do not involve any
special treatment regarding the semantics of related actions.
\modLP{%
It is also worth noticing that the use of such intermediate sorts may trigger
spurious transitions due to potential incoherences between the state of the
cooperative sort and the actual state of cooperating sorts.
However, such phenomena can be fixed by the use of priorities between actions \cite{FPMR13-CS2Bio}.}
